# Import the spreadsheets: series data

## read data

For this scale of data, I'd rather use `read_xlsx()` instead of `googlesheets4::read_sheet()`

```{r, message=FALSE, warning=FALSE}
excel_path <- "./input/Pokemon TCG Spreadsheet V3.25.xlsx"
sheet_names <- excel_sheets(excel_path)
sheet_names2 <- sheet_names[-1] # remove "Overview" sheet

release_dates <- 
  map(
    sheet_names2,
    ~ read_xlsx(
      excel_path,
      sheet = .x, 
      range = "B1:B2"
    )
  ) |> 
  map(~ pull(.x)) |> 
  reduce(c)
release_dates_df <- 
  tibble(
    series = sheet_names2, 
    release_date = release_dates
  )
release_dates_df
overview_series_range <- c(
  "D4:D12",
  "G4:G11",
  "J4:J26",
  "M4:M22",
  "D31:D43",
  "G31:G50",
  "J31:J47",
  "M31:M46"
  )
series_gen <- 
  map(
    .x = overview_series_range,
    ~ read_xlsx(
      excel_path,
      col_names = "series",
      sheet = "Overview", 
      range = .x
    )
  ) |> 
  map2(
    .y = 1:8,
    ~ mutate(.data = .x, series_gen = rep(.y, nrow(.x)))
  ) |> 
  bind_rows()
release_dates_df2 <- 
  left_join(release_dates_df, series_gen, by = "series") |> 
  arrange(release_date) |> 
  mutate(
    lead = lead(series_gen),
    series_gen = if_else(is.na(series_gen), lead, series_gen),
    lead = NULL) 
release_dates_df2
```

## release date, main expansion or not, from bulbagarden

Use [Convert Wiki Tables to CSV](https://wikitable2csv.ggor.de/).
Set the Table Selector to `.sortable` and fetch tables provided by [bulbagarden](https://bulbapedia.bulbagarden.net/wiki/List_of_Pok%C3%A9mon_Trading_Card_Game_expansions)

|csv no.|content|name|
|---|---|---|
|1|Main expansions|`List_of_Pok%C3%A9mon_Trading_Card_Game_expansions_1.csv`|
|2|Special expansions|`List_of_Pok%C3%A9mon_Trading_Card_Game_expansions_2.csv`|
|3|Black Star Promos|`List_of_Pok%C3%A9mon_Trading_Card_Game_expansions_3.csv`|
|4|POP Series|`List_of_Pok%C3%A9mon_Trading_Card_Game_expansions_4.csv`|
|5|McDonald's Collections|`List_of_Pok%C3%A9mon_Trading_Card_Game_expansions_5.csv`|


```{r}
series_main_expansions_df <- read_csv(
  file = "./input/List_of_Pok%C3%A9mon_Trading_Card_Game_expansions_1.csv"
) |> 
  mutate(series_class = "Main Expansion")
colnames(series_main_expansions_df)
series_special_expansions_df <- read_csv(
  file = "./input/List_of_Pok%C3%A9mon_Trading_Card_Game_expansions_2.csv"
) |> 
  mutate(series_class = "Special Expansion")
series_black_star_promos_df <- read_csv(
  file = "./input/List_of_Pok%C3%A9mon_Trading_Card_Game_expansions_3.csv"
) |> 
  mutate(series_class = "Black Star Promo")
series_pop_series_df <- read_csv(
  file = "./input/List_of_Pok%C3%A9mon_Trading_Card_Game_expansions_4.csv"
) |> 
  mutate(series_class = "Pop Series")
series_mcdonalds_df <- read_csv(
  file = "./input/List_of_Pok%C3%A9mon_Trading_Card_Game_expansions_5.csv"
) |> 
  mutate(series_class = "McDonalds Collection",
         `EN cards` = as.character(`EN cards`))
series_mcdonalds_df
```


```{r}
series_classification_df <- bind_rows(
  series_main_expansions_df, 
  series_special_expansions_df, 
  series_black_star_promos_df,
  series_pop_series_df,
  series_mcdonalds_df)
series_classification_df
```

## merge with release dates df

Merge `series_classification_df ` with `release_dates_df2`
```{r}
series_classification_df |> colnames()
release_dates_df2 |> colnames()
# str_view_all("October, 2003 - September, 2006 ", ".+(?=\\s-)")
# str_view_all("September 7, 2018 ", "\\s\\*\\s|(?<=[a-z]{3,10}\\s\\d{4})\\s")
# 
# str_view_all(
#   "January, 2003 - February, 2004 February, 2004 - July, 2006", 
#   "\\s\\*\\s|(?<=[a-z]{3,13},?\\s\\d{4})\\s(?!-)")
```


```{r}
series_classification_df2 <- 
  series_classification_df |> 
  rename(
    series = `English name`,
    series_ja = `Japanese name`,
    series_abb = `Set abb.`,
    cards_total = `EN cards`,
    cards_total_ja = `JP cards`,
    release_date = `EN release date`,
    release_date_ja = `JP release date`,
         ) |> 
  select(-`EN symbol`, -`JP symbol`, -`EN no.`, -`JP no.`) |>
  # separate https://tidyr.tidyverse.org/reference/separate.html
  separate(
    cards_total_ja, 
    into = c("cards_total_ja_1", "cards_total_ja_2"),
    fill = "right"
  )
series_classification_df2
```


```{r}
# separate release dates of Japanese versions that are released as separate series
series_classification_df3 <- 
  series_classification_df2 |> 
  mutate(
    release_date_ja = 
      str_replace(
        release_date_ja,
        "\\s\\*\\s|(?<=[a-z]{3,13},?\\s\\d{4})\\s(?!-)",
        "SEPARATE"
      )) |> 
  separate(
    release_date_ja, 
    into = c("release_date_ja_1", "release_date_ja_2"),
    sep = "\\s\\*\\s|SEPARATE",
    fill = "right"
  ) |> 
  mutate(
    across(
      everything(), 
      ~ recode(
        .,
        "—" = NA_character_, 
        "Ongoing" = NA_character_)
    ),
    release_date_ja_2 = str_replace(release_date_ja_2, "\\s\\*", "")
  )
series_classification_df3
```


```{r}
# test
# str_view_all(
#   "From February 9, 2021", 
#   "January|February|March|April|May|June|July|August|September|October|November|December")
# str_view_all(
#   "April - August, 2006", 
#   "[:digit:]{4}")
# str_view_all(
#   "October 13th - November 26th, 2013", 
#   "[:digit:]{4}")
# str_view_all(
#   "September, 2008 - March, 2009", 
#   "[:digit:]{4}")

# https://lubridate.tidyverse.org/reference/parse_date_time.html

parse_date_time(
  x = "2021 February 21",
  orders = "ybd"
)

# run

series_classification_df4 <-
  series_classification_df3 |> 
  mutate(
    release_date_month = 
      str_extract(
        release_date,
        "January|February|March|April|May|June|July|August|September|October|November|December"
      ),
    release_date_day = 
      str_extract(
        release_date,
        "[:digit:]{1,3}(?=,)|[:digit:]{1,3}(?=[:lower:]{2})(?![:lower:]{2},)"
      ) |> replace_na(replace = "1"),
    release_date_year = 
      str_extract(
        release_date,
        "[:digit:]{4}"
      ),
    release_date = 
      str_c(
        release_date_year, 
        release_date_month, 
        release_date_day
      ) |> 
      parse_date_time(orders = "ybd"),
    release_date_month = 
      str_extract(
        release_date_ja_1,
        "January|February|March|April|May|June|July|August|September|October|November|December"
      ),
    release_date_day = 
      str_extract(
        release_date_ja_1,
        "[:digit:]{1,3}(?=,)|[:digit:]{1,3}(?=[:lower:]{2})(?![:lower:]{2},)"
      ) |> replace_na(replace = "1"),
    release_date_year = 
      str_extract(
        release_date_ja_1,
        "[:digit:]{4}"
      ),
    release_date_ja_1 = 
      str_c(
        release_date_year, 
        release_date_month, 
        release_date_day
      ) |> 
      parse_date_time(orders = "ybd"),
    release_date_month = 
      str_extract(
        release_date_ja_2,
        "January|February|March|April|May|June|July|August|September|October|November|December"
      ),
    release_date_day = 
      str_extract(
        release_date_ja_2,
        "[:digit:]{1,3}(?=,)|[:digit:]{1,3}(?=[:lower:]{2})(?![:lower:]{2},)"
      ) |> replace_na(replace = "1"),
    release_date_year = 
      str_extract(
        release_date_ja_2,
        "[:digit:]{4}"
      ),
    release_date_ja_2 = 
      str_c(
        release_date_year, 
        release_date_month, 
        release_date_day
      ) |> 
      parse_date_time(orders = "ybd"),
  )
series_classification_df4
```


```{r}
series_classification_df5 <- 
  series_classification_df4 |> 
  select(
    series_class,
    series,
    series_abb,
    series_ja,
    cards_total,
    release_date,
    cards_total_ja_1,
    cards_total_ja_2,
    release_date_ja_1,
    release_date_ja_2,
  )
series_classification_df5
```

[Multiple observations per row](https://tidyr.tidyverse.org/articles/pivot.html#multiple-observations-per-row)を参考に、`cards_total_ja`と`release_date_ja`を一気にpivot_longerすることが…できるのでは・・・
```{r}
series_classification_df6 <-
  series_classification_df5 |> 
  # rowid_to_column("id") |>
  pivot_longer(
    !series_class:release_date,
    # !series_class:release_date, 
    names_to = c(".value", "ja_variation"),
    names_pattern = "(.*)_(1|2)",
    # names_sep = "_", 
    values_drop_na = TRUE
  ) |> 
  select(-ja_variation)
series_classification_df6
```

## join

表記ゆれや数日程度の`release_date`のずれがある。基本的にBulbapediaのものにあわせる
sheet名もそれに伴って書き換えなければいけないが、仕方ない
```{r}
release_dates_df3 <-
  release_dates_df2 |> 
  mutate(
    series = recode(
      series,
      `McDonalds Collection 2021` = "McDonald's Collection 2021",
      `Champions Path` = "Champion's Path",
      `Sword and Shield` = "Sword & Shield",
      `Sword and Shield Promos` = "SWSH Black Star Promos",
      `Sun and Moon` = "Sun & Moon",
      `Sun and Moon Promos` = "SM Black Star Promos",
      `Mcdonalds Collection 2015` = "McDonald's Collection 2015",
      `Mcdonalds Collection 2014` = "McDonald's Collection 2014",
      `McDonalds Collection 2012` = "McDonald's Collection 2012", # not Mcd!
      `X and Y` = "XY",
      `X and Y Promos` = "XY Black Star Promos",
      `McDonalds Collection` = "McDonald's Collection",
      `Black and White` = "Black & White", 
      `HeartGold and SoulSilver` = "HeartGold & SoulSilver"
    ),
    series = str_replace(series, "Pop", "POP")
  )
release_dates_df3
```
つぎに、Bulbapediaを母体に一緒にしつつもTrainer KitはBulbapediaのリストにはないため`series_class`に`Trainer Kit`を新設しつつ、`release_date`についてはBulbapediaを優先する。
```{r}
series_release_df1 <- full_join(
  series_classification_df6, 
  release_dates_df3,
  by = c("series")) |> 
  distinct() |> 
  mutate(
    series_class = 
      case_when(
        str_detect(series, "Trainer\\sKit") ~ "Trainer Kit", 
        TRUE ~ series_class
      )
  )
series_release_df1
```


```{r}
df_series <- series_release_df1 |> 
  # select(release_date.x, release_date.y) |> 
  mutate(
    # indicate if the series is in V3.25 spreadsheet or 
    # only available on bulbagarden's list
    meta_is_bulba_only = is.na(release_date.y),
    meta_is_v325_only = series_class == "Trainer Kit",
    release_date = 
      case_when(
        is.na(release_date.x) & !is.na(release_date.y) ~ release_date.y,
        
        !is.na(release_date.x) & is.na(release_date.y) ~ release_date.x,
        release_date.x == release_date.y ~ release_date.x,
        release_date.x != release_date.y ~ release_date.x,
        # TRUE ~ NA_Date_
      ),
      release_date.x = NULL,
      release_date.y = NULL,
      cards_total = cards_total |> as.character() |> as.integer(),
      cards_total_ja = cards_total |> as.character() |> as.integer(),
      series_class = case_when(
        series == "Pokemon GO" ~ "Special Expansion",
        TRUE ~ series_class
      ),
    # across(starts_with("release_date"), ~ as_date) nazeka ugokan
    release_date = release_date |> as_date(),
    release_date_ja = release_date_ja |> as_date()
  ) |> 
  select(series_class, series, release_date, cards_total, series_gen, series_abb,  series_ja, release_date_ja, cards_total_ja, everything() )
df_series # 157 x 11
```

## df_series: colnames explanation

```{r}
df_series |> colnames()

```

|`colname`|`meaning`|str|
|---|---|---|
|`series_class`|Retrieved from Bulbagarden. Either `"Main Expansion"`, `"Special Expansion"`, `"Black Star Promo"`, `"Pop Series"`, `"McDonalds Collection"`, or `"Trainer Kit"`.|chr|
| `series`|The name of the series. "`Neo Genesis`", "`Sun & Moon`" etc. The names are based on the ones listed in Bulbapedia, not the [V3.25 spreadsheet](https://docs.google.com/spreadsheets/d/10G8wEY70qJ7rEnGbDzJp13vdOB4ikIjz4E3SQVRglRc/edit#gid=59182613).|chr|
|`release_date`|The release date of the series. The dates are based on the ones listed in Bulbapedia, not the ones on the [V3.25 spreadsheet](https://docs.google.com/spreadsheets/d/10G8wEY70qJ7rEnGbDzJp13vdOB4ikIjz4E3SQVRglRc/edit#gid=59182613), from which the Bulbapedia's date slightly differ. Release dates with no specific dates are converted to the 1st day of the month, e.g., `May 2003` -> `2003-05-01`.|Date|
|`cards_total`|the amount of total cards of the series, according to Bulbapedia. The official count is sometimes lower than this. This does not necessarily match the amount of total cards of the series that can be calculated from the card list. The card list is based on V3.25 spreadsheet and the data is not exhaustive.|int|
|`series_gen`|generation of the series, from 1-8. This is based on V3.25 spreadsheet and may not match from Bulbapedia's classification (e.g., `Legendary Collection` classified as series_gen 1 in spreadsheet, series_gen 2 in Bulbapedia).|int|
|`series_abb`|from Bulbapedia. might be useful for plotting to avoid overcrowding.|chr|
|`series_ja`|Name of the series in Japanese version according to Bulbapedia.|chr|
|`release_date_ja`||Date|
|`cards_total_ja`||int|
|`meta_is_bulba_only`|`TRUE` if the series is only from Bulbapedia.|lgl|
|`meta_is_v325_only`|`TRUE` if the series is only from spreadsheet. "Trainer Kit"s.|lgl|

In this chapter we tidied the data for the series (`df_series`). In the next chapter `pokemon-cards` we tidy the data for the individual cards (`df`).

