# df_names

[Bulbapedia](https://bulbapedia.bulbagarden.net/wiki/List_of_Japanese_Pok%C3%A9mon_names) once again to produce [multiple csvs](https://stackoverflow.com/questions/11433432/how-to-import-multiple-csv-files-at-once). The CSVs are stored under `./input/bulbapedia_name`.
```{r}
files <- list.files(
  path = "./input/bulbapedia_name", 
  pattern = "*.csv",
  full.names = TRUE)
names_df <- map(
  files,
  ~ read_csv(
    file = .x, 
    skip = 1, 
    col_types = "c"
  ) 
) |> 
  map2(
    .y = 1:9,
    ~ mutate(.data = .x, pokemon_gen = rep(.y, nrow(.x)))
  ) |> 
  bind_rows() |> 
  rename(
    pokedex_id = Ndex,
    pokemon_name = English,
    pokemon_name_ja = Kana
    ) |> 
  select(pokedex_id, pokemon_name, pokemon_name_ja, pokemon_gen)
names_df2 <- names_df |> 
  mutate(
    pokedex_id = pokedex_id |> as.character() |> as.integer(),
  )
```

## images

large images (`image_large`, in `./input/image_kaggle/`) are from [kaggle](https://www.kaggle.com/datasets/arenagrenade/the-complete-pokemon-images-data-set?resource=download), and missing images (mainly new ones around pokedex_id 900) downloaded manually from official pokedex sarch.
small images (`image_small`, in `./input/bulbapedia_image/`) are from [bulbapedia](https://bulbapedia.bulbagarden.net/wiki/List_of_Pok%C3%A9mon_by_National_Pok%C3%A9dex_number).
use cmd+i > media to download. it fails for some of the files though... :(

tips: create [leading zeros](https://stackoverflow.com/questions/8266915/format-number-as-fixed-width-with-leading-zeros)


```{r, echo = FALSE}
image_file_names <- list.files("./input/image_kaggle", pattern = "png")
# new_names <- image_file_names |> str_remove("70px-") |> str_remove("(?<=[:digit:]{3}).+(?=\\.)")
new_names <-
  image_file_names %>% 
  if_else(
    str_detect(., pattern = "mime|mr-|hakamo|ho-oh|-o\\.|porygon-z|^tapu-|^type-"), 
    ., 
    str_remove(., pattern = "-.+(?=\\.png)")
  ) 
file.rename(
  paste0("./input/image_kaggle/",image_file_names),
  to = paste0("./input/image_kaggle/", new_names)
) |> invisible()
```


```{r}
df_names <-
  names_df2 |> 
  filter(!is.na(pokedex_id)) |>  # remove gen 9 pokemons
  mutate(
    pokemon_name_simplified = tolower(pokemon_name) |> 
      str_replace("\\. ", "-") |> 
      str_remove("\\.") |> 
      str_remove("\\'") |> 
      str_replace("mime jr", "mime-jr") |> 
      str_replace("tapu ", "tapu-") |> 
      str_replace("type: null", "type-null") |> 
      stri_trans_general("Latin-ASCII"), 
    image_large = paste0("./input/image_kaggle/", pokemon_name_simplified, ".png"),
    image_small = paste0(
    "./input/bulbapedia_image/",
    sprintf("%03d", pokedex_id),
    # pokemon_name,
    ".png"
    )
  )
```

### plot all 905 pokemons
```{r}
# df_names |> 
#   # filter(pokedex_id < 100) |> 
#   ggplot(
#     aes(x = pokedex_id %/% 30, y = pokedex_id %% 30, image = image_small)) +
#   geom_image(size = .04) +
#   theme_pokemon +
#   theme(
#     aspect.ratio = 1
#   )
# ggsave("./output/pokemons.png", width = 5000, height = 5000, units = "px")

```

## pokemon sequencing

```{r}
df_gene_pokemon <- df_card |> # from 02 
  rowid_to_column(var = "card_id") |> 
  filter(is_pokemon) |>
  mutate(
    nm = card_name
  ) |> 
  mutate(
    is_gx = 
      str_detect(
        nm,
        "gx$|GX$"
      ),
    nm = # remove GX from name
      str_replace(
        nm,
        "GX$",
        ""
      ),
    nm = # remove "L" from legend team poke's and them only
      str_replace( 
      nm,
      "(?<=\\&\\s[:alpha:]{3,13})L$",
      ""
    ),
    is_team = 
      str_detect(
        nm,
        "\\s\\&\\s"),
    is_tag_team = is_team & is_gx,
    # though not properly named, Entei & Raikou is Legend
    # https://bulbapedia.bulbagarden.net/wiki/Entei_%26_Raikou_LEGEND_(Unleashed_90)
    is_legend = is_team & !is_gx, 
    
    ) |>
  separate(
    col = nm,
    sep = "\\s\\&\\s",
    into = c("nm", "team_buddy1", "team_buddy2")
           ) |> 
  mutate(
    is_extra = str_detect(nm, "\\sex$|\\sex"),
    nm = str_replace(nm, "\\sex$|\\sex", ""),
    is_ex = str_detect(nm, "EX$"),
    nm = str_replace(nm, "EX$", ""),
    has_master = str_detect(nm, "^.+\\'s\\s"),
    master = str_extract(nm, "^.+(?=\\'s\\s)"),
    nm = str_replace(nm, "^.+\\'s\\s", ""),
    is_dark = str_detect(nm, "Dark\\s"),
    nm = str_replace(nm, "Dark\\s", ""),
    is_delta = str_detect(nm, "\\sδ"), 
    nm = str_replace(nm, "\\s{0,3}δ", ""),
    is_light = str_detect(nm, "^Light\\s"),
    nm = str_replace(nm, "^Light\\s", ""),
    is_star = str_detect(nm, "\\*$"),
    nm = str_replace(nm, "\\*$", ""),
    is_bros = str_detect(nm, "\\sBros\\.$"),
    nm = str_replace(nm, "\\sBros\\.$", ""),
    is_levelten = str_detect(nm, "\\s?LV\\.X"),
    nm = str_replace(nm, "\\s?LV\\.X", ""),
    is_ginga = str_detect(nm, "\\s?G$"),
    nm = str_replace(nm, "\\s?G$", ""),
    is_gymleader = str_detect(nm, "\\s?GL$"),
    nm = str_replace(nm, "\\s?GL$", ""),
    is_elitefour = str_detect(nm, "\\s?4$"),
    nm = str_replace(nm, "\\s?4$", ""),
    is_champion = str_detect(nm, "C$"),
    nm = str_replace(nm, "\\s?C$", ""),
    is_frontierbrain = str_detect(nm, "\\s?FB$"),
    nm = str_replace(nm, "\\s?FB$", ""),
    is_break = str_detect(nm, "\\s?BREAK"),
    nm = str_replace(nm, "\\s?BREAK", ""),
    is_primal = str_detect(nm, "^Primal\\s"),
    nm = str_replace(nm, "^Primal\\s", ""),
    is_alolan = str_detect(nm, "^Alolan\\s"),
    nm = str_replace(nm, "^Alolan\\s", ""),
    is_galarian = str_detect(nm, "^Galarian\\s"),
    nm = str_replace(nm, "^Galarian\\s", ""),
    is_hisuian = str_detect(nm, "^Hisuian\\s"),
    nm = str_replace(nm, "^Hisuian\\s", ""),
    is_v_family = str_detect(nm, "V$|VMAX$|VUNION$|VSTAR$"),
    is_v = str_detect(nm, "(?<=[:lower:])V$"),
    nm = str_replace(nm, "(?<=[:lower:])V$", ""), # avoid "Unown V"
    is_vmax = str_detect(nm, "VMAX$"),
    nm = str_replace(nm, "VMAX$", ""),
    is_vunion = str_detect(nm, "VUNION$"),
    nm = str_replace(nm, "VUNION$", ""),
    is_vstar = str_detect(nm, "VSTAR$"),
    nm = str_replace(nm, "VSTAR$", ""),
    is_prismstar = str_detect(nm, "♦$"),
    nm = str_replace(nm, "♦$", ""),
    is_mega = str_detect(nm, "^M(?=[:upper:])|^Mega\\s"),
    nm = str_replace(nm, "^M(?=[:upper:])|^Mega\\s", ""),
    
    is_originform = str_detect(nm, "^Origin\\sForme\\s"),
    nm = str_replace(nm, "^Origin\\sForme\\s", ""),
    is_shining = str_detect(nm, "^Shining\\s"),
    nm = str_replace(nm, "^Shining\\s", ""),
    is_specialdelivery = str_detect(nm, "^Special\\sDelivery\\s"),
    nm = str_remove(nm, "^Special\\sDelivery\\s"),
    # LEGEND, L and LEGEND 
    is_legend = is_legend | str_detect(nm, "LEGEND$"),
    nm = str_replace(nm, "LEGEND$", ""),
    
    # extract formes
    # kyurem_colour = 
    # Unown includes ! and 
    forme_variant = 
      str_extract(
        nm, 
        "(?<=Shellos).+|(?<=Gastrodon).+|(?<=Wormadam).+|(?<=Burmy).+|(?<=Unown).+|(?<=Castform).+|.+(?=Castform)|(?<=Ditto).+|(?<=Deoxys).+|.+(?=Necrozma$)|.+(?=Porygon$)|.+(?=Greninja$)|.+(?=Mewtwo$)|.+(?=Urshifu$)|.+(?=Calyrex$)|.+(?=Kyurem)|.+(?=Rotom)|Detective\\s|Radiant\\s|\\sLibre|Sunny\\s|Surfing\\s|Flying\\s") |> 
      str_remove("^\\s|\\s$"), # remove spaces before and after formes
    nm =
      str_remove(
        nm, 
        "(?<=Shellos).+|(?<=Gastrodon).+|(?<=Wormadam).+|(?<=Burmy).+|(?<=Unown).+|(?<=Castform).+|.+(?=Castform)|(?<=Ditto).+|(?<=Deoxys).+|.+(?=Necrozma$)|.+(?=Porygon$)|.+(?=Greninja$)|.+(?=Mewtwo$)|.+(?=Urshifu$)|.+(?=Calyrex$)|.+(?=Kyurem)|.+(?=Rotom)|Detective\\s|Radiant\\s|\\sLibre|Sunny\\s|Surfing\\s|Flying\\s"
    ),
  ) |> 
  left_join(df_names, by = c("nm" = "pokemon_name")) |> 
  rename(pokemon_name = nm) |> 
  rowwise() |> 
  mutate(
    pokemon_deco = str_remove(
      card_name, 
      pattern = paste(pokemon_name)
    )
  ) |> 
  select(
    card_id, 
    card_name, 
    pokedex_id, 
    pokemon_name,
    pokemon_name_ja, 
    pokemon_deco, 
    pokemon_gen, 
    card_type2,
    team_buddy1, 
    team_buddy2, 
    set, 
    series,
    release_date, 
    everything(),
    starts_with("is_"), 
    starts_with("has_"), 
  ) |> 
  ungroup()
```
warnings

> Warning: Expected 3 pieces. Missing pieces filled with `NA` in 12713 rows [1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, ...].

are for GX tag team buddies, ignore them.

## remove pokemon name from original name to extract subgenes

subgene or subpokemon or parasites or decorations or whatever

```{r}
df_gene_pokemon # 12325 x 56
df_gene <- bind_rows(
  df_gene_pokemon,
  df_card |> rowid_to_column(var = "card_id") |> filter(!is_pokemon)
  ) |>
  arrange(card_id)
  
```
### summary of the data
紀要掲載予定

```{r}
df_gene |> count() # 14565
df_gene |> count(is_pokemon) # 12325, non-2240
df_gene |> 
  group_by(series) |> 
  count() # 125 series
df_gene |> 
  filter(series_class != "Black Star Promo") |> 
  filter(release_date == max(release_date) )

```

