# df_names

Create a list of the most basic names of pokemons, e.g. "pikachu", along with their pokedex # and generation. Kaggle's data wasn't good, [Github](https://github.com/lgreski/pokemonData/blob/master/Pokemon.csv)'s data wasn't either.

You don't have to run the next chunk:

```{r}
# pokemon_names_df_raw <- read_csv("./input/pokemon_github.csv")
# pokemon_names_df_raw_ja <- read_csv("./input/ポケモンデータシート.csv")
# pokemon_names_df_raw_ja |> colnames()
# pokemon_names_df1 <- pokemon_names_df_raw |> 
#   # filter(Form %in% c("Normal Forme", NA)) |> 
#   select(
#     ID, card_name, Generation) |> 
#   rename(
#     id = ID,
#     name = card_name, 
#     pokemon_gen = Generation
#          ) |> 
#   distinct() |> 
#   filter(!str_detect(name, "50"))
# pokemon_names_df1 |> dim()
# pokemon_names_df_ja <- pokemon_names_df_raw_ja |> 
#   rename(
#     id = `ぜんこくNo.`,
#     name_ja = `名前`,
#     name = `英語名`
#   ) |> 
#   select(id, name_ja, name)
# df_names <- 
#   full_join(pokemon_names_df1, pokemon_names_df_ja, by = c("name", "id")) |> 
#   distinct() |> 
#   mutate(id = id |> as.character() |> as.integer()) |> 
#   select(id, name, name_ja, pokemon_gen)
# df_names |> View()
# pokemon_names_df_ja |> 
#   filter(id == "469")
#  
#   View(pokemon_names_df1 )
```
bad bad bad. I resort to 
[Bulbapedia](https://bulbapedia.bulbagarden.net/wiki/List_of_Japanese_Pok%C3%A9mon_names) once again to produce [multiple csvs](https://stackoverflow.com/questions/11433432/how-to-import-multiple-csv-files-at-once). The CSVs are stored under `./input/bulbapedia_name`.
```{r}
files <- list.files(
  path = "./input/bulbapedia_name", 
  pattern = "*.csv",
  full.names = TRUE)
names_df <- map(
  files,
  ~ read_csv(
    file = .x, 
    skip = 1, 
    col_types = "c"
  ) 
) |> 
  map2(
    .y = 1:9,
    ~ mutate(.data = .x, pokemon_gen = rep(.y, nrow(.x)))
  ) |> 
  bind_rows() |> 
  rename(
    pokedex_id = Ndex,
    name = English,
    name_ja = Kana
    ) |> 
  select(pokedex_id, name, name_ja, pokemon_gen)
names_df2 <- names_df |> 
  mutate(
    pokedex_id = pokedex_id |> as.character() |> as.integer(),
  )
df_names <- names_df2 |> 
  filter(!is.na(pokedex_id)) # remove gen 9 pokemons
```

## pokemon sequencing

```{r}
df_gene <- df_card |> # from 02 
  filter(is_pokemon) |> 
  mutate(
    nm = card_name
  ) |> 
  mutate(
    is_gx = 
      str_detect(
        nm,
        "gx$|GX$"
      ),
    nm = # remove GX from name
      str_replace(
        nm,
        "GX$",
        ""
      ),
    nm = # remove "L" from legend team poke's and them only
      str_replace( 
      nm,
      "(?<=\\&\\s[:alpha:]{3,13})L$",
      ""
    ),
    is_team = 
      str_detect(
        nm,
        "\\s\\&\\s"),
    is_tag_team = is_team & is_gx,
    # though not properly named, Entei & Raikou is Legend
    # https://bulbapedia.bulbagarden.net/wiki/Entei_%26_Raikou_LEGEND_(Unleashed_90)
    is_legend = is_team & !is_gx, 
    
    ) |>
  separate(
    col = nm,
    sep = "\\s\\&\\s",
    into = c("nm", "team_buddy1", "team_buddy2")
           ) |> 
  mutate(
    is_extra = str_detect(nm, "\\sex$|\\sex"),
    nm = str_replace(nm, "\\sex$|\\sex", ""),
    is_ex = str_detect(nm, "EX$"),
    nm = str_replace(nm, "EX$", ""),
    has_master = str_detect(nm, "^.+\\'s\\s"),
    master = str_extract(nm, "^.+(?=\\'s\\s)"),
    nm = str_replace(nm, "^.+\\'s\\s", ""),
    is_dark = str_detect(nm, "Dark\\s"),
    nm = str_replace(nm, "Dark\\s", ""),
    is_delta = str_detect(nm, "\\sδ"), 
    nm = str_replace(nm, "\\s{0,3}δ", ""),
    is_light = str_detect(nm, "^Light\\s"),
    nm = str_replace(nm, "^Light\\s", ""),
    is_star = str_detect(nm, "\\*$"),
    nm = str_replace(nm, "\\*$", ""),
    is_bros = str_detect(nm, "\\sBros\\.$"),
    nm = str_replace(nm, "\\sBros\\.$", ""),
    is_levelten = str_detect(nm, "\\s?LV\\.X"),
    nm = str_replace(nm, "\\s?LV\\.X", ""),
    is_ginga = str_detect(nm, "\\s?G$"),
    nm = str_replace(nm, "\\s?G$", ""),
    is_gymleader = str_detect(nm, "\\s?GL$"),
    nm = str_replace(nm, "\\s?GL$", ""),
    is_elitefour = str_detect(nm, "\\s?4$"),
    nm = str_replace(nm, "\\s?4$", ""),
    is_champion = str_detect(nm, "C$"),
    nm = str_replace(nm, "\\s?C$", ""),
    is_frontierbrain = str_detect(nm, "\\s?FB$"),
    nm = str_replace(nm, "\\s?FB$", ""),
    is_break = str_detect(nm, "\\s?BREAK"),
    nm = str_replace(nm, "\\s?BREAK", ""),
    is_primal = str_detect(nm, "^Primal\\s"),
    nm = str_replace(nm, "^Primal\\s", ""),
    is_alolan = str_detect(nm, "^Alolan\\s"),
    nm = str_replace(nm, "^Alolan\\s", ""),
    is_galarian = str_detect(nm, "^Galarian\\s"),
    nm = str_replace(nm, "^Galarian\\s", ""),
    is_hisuian = str_detect(nm, "^Hisuian\\s"),
    nm = str_replace(nm, "^Hisuian\\s", ""),
    is_v_family = str_detect(nm, "V$|VMAX$|VUNION$|VSTAR$"),
    is_v = str_detect(nm, "(?<=[:lower:])V$"),
    nm = str_replace(nm, "(?<=[:lower:])V$", ""), # avoid "Unown V"
    is_vmax = str_detect(nm, "VMAX$"),
    nm = str_replace(nm, "VMAX$", ""),
    is_vunion = str_detect(nm, "VUNION$"),
    nm = str_replace(nm, "VUNION$", ""),
    is_vstar = str_detect(nm, "VSTAR$"),
    nm = str_replace(nm, "VSTAR$", ""),
    is_prismstar = str_detect(nm, "♦$"),
    nm = str_replace(nm, "♦$", ""),
    is_mega = str_detect(nm, "^M(?=[:upper:])|^Mega\\s"),
    nm = str_replace(nm, "^M(?=[:upper:])|^Mega\\s", ""),
    
    is_originform = str_detect(nm, "^Origin\\sForme\\s"),
    nm = str_replace(nm, "^Origin\\sForme\\s", ""),
    is_shining = str_detect(nm, "^Shining\\s"),
    nm = str_replace(nm, "^Shining\\s", ""),
    is_specialdelivery = str_detect(nm, "^Special\\sDelivery\\s"),
    nm = str_remove(nm, "^Special\\sDelivery\\s"),
    # LEGEND, L and LEGEND 
    is_legend = is_legend | str_detect(nm, "LEGEND$"),
    nm = str_replace(nm, "LEGEND$", ""),
    
    # extract formes
    # kyurem_colour = 
    # Unown includes ! and 
    forme_variant = 
      str_extract(
        nm, 
        "(?<=Shellos).+|(?<=Gastrodon).+|(?<=Wormadam).+|(?<=Burmy).+|(?<=Unown).+|(?<=Castform).+|.+(?=Castform)|(?<=Ditto).+|(?<=Deoxys).+|.+(?=Necrozma$)|.+(?=Porygon$)|.+(?=Greninja$)|.+(?=Mewtwo$)|.+(?=Urshifu$)|.+(?=Calyrex$)|.+(?=Kyurem)|.+(?=Rotom)|Detective\\s|Radiant\\s|\\sLibre|Sunny\\s|Surfing\\s|Flying\\s") |> 
      str_remove("^\\s|\\s$"), # remove spaces before and after formes
    nm =
      str_remove(
        nm, 
        "(?<=Shellos).+|(?<=Gastrodon).+|(?<=Wormadam).+|(?<=Burmy).+|(?<=Unown).+|(?<=Castform).+|.+(?=Castform)|(?<=Ditto).+|(?<=Deoxys).+|.+(?=Necrozma$)|.+(?=Porygon$)|.+(?=Greninja$)|.+(?=Mewtwo$)|.+(?=Urshifu$)|.+(?=Calyrex$)|.+(?=Kyurem)|.+(?=Rotom)|Detective\\s|Radiant\\s|\\sLibre|Sunny\\s|Surfing\\s|Flying\\s"
    ),
  ) |> 
  left_join(df_names, by = c("nm" = "name")) |> 
  rename(pokemon_name = nm)
```
warnings

> Warning: Expected 3 pieces. Missing pieces filled with `NA` in 12713 rows [1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, ...].

are for GX tag team buddies, ignore them.

## remove pokemon name from original name to extract subgenes

subgene or subpokemon or parasites or decorations or whatever

```{r}
df_gene |> colnames()
df_gene <- df_gene |>   
  rowwise() |> 
  mutate(pokemon_deco = str_remove(card_name, pattern = paste(pokemon_name))) |> 
  select(pokedex_id, card_name, pokemon_name, name_ja, pokemon_deco, pokemon_gen, card_type2, team_buddy1, team_buddy2, set, series, release_date, everything(), starts_with("is_"), starts_with("has_")) |> 
  ungroup()
```
