# individual poke

最も単純な帰無モデルと比較するために、リアルデータのCCDFを両対数グラフで書く。

## まずはなにもせずカウント
```{r}
df_poke_count_raw <- df_gene |> 
  filter(is_pokemon) |> 
  group_by(card_name) |> 
  summarise(n = n()) |> 
  arrange(desc(n))

```

### ECDF

[負うた子に教えられて浅瀬を渡る](https://qiita.com/xerroxcopy/items/b79635ef3dbcc29644c6)

```{r}
# pokemon name. the only datatype that is comparable to the random draw null model

df_count_pokemon <- df_gene |> 
  filter(is_pokemon) |> 
  group_by(pokemon_name) |> 
  summarise(n = n()) |> 
  arrange(desc(n)) |> 
  mutate(ecdf = 1 - ecdf(n)(n - .01)) |> 
  select(-pokemon_name) |> 
  distinct() |> 
  mutate(condition = "pokémon name") |> 
  arrange(desc(n))


# card name. 
df_count_full_name <- df_gene |> 
  filter(is_pokemon) |> 
  group_by(card_name) |> 
  summarise(n = n()) |> 
  arrange(desc(n)) |> 
  mutate(ecdf = 1 - ecdf(n)(n - .01)) |>
  select(-card_name) |> 
  distinct() |> 
  mutate(condition = "card name")


# card name + type. 
# distinguish pokemons with the same name, but with different types
df_count_full_name_type <- df_gene |> 
  filter(is_pokemon) |> 
  mutate(name_type = paste(card_name, card_type2)) |> 
  group_by(name_type) |> 
  summarise(n = n()) |> 
  arrange(desc(n)) |> 
  mutate(ecdf = 1 - ecdf(n)(n - .01)) |>
  select(-name_type) |> 
  distinct() |> 
  mutate(condition = "card name + type")

# advanced:
# full name, but remove deco tsuki cards
# count only the cards named "Pikachu" for pikachu.
df_count_full_name_remove_decorated <- df_gene |> 
  filter(is_pokemon) |> 
  filter(card_name == pokemon_name) |> 
  group_by(card_name) |> # same as grouping by pokemon_name since card_name == pokemon_name
  summarise(n = n()) |> 
  arrange(desc(n)) |> 
  mutate(ecdf = 1 - ecdf(n)(n - .01)) |>
  select(-card_name) |> 
  distinct() |> 
  mutate(condition = "card name, remove decorated")
```

`ecdf()`はX > nである割合であってX >= nである割合ではないのだな
やりたいこと的には、`n = 1`のときに、1以上のコピーがあるポケカの割合は1になってないとおかしい。n = 1のときに1よりもたくさんのコピーがある割合1 - .32とかを計算してしまっている。同様にn = 58のとき（ピカチュウさん）、X >= 58である割合は 1 / 2425 (2425はユニークなポケモン名の数、`df$card_name |> unique() |> length()`)かと思うが、0になってしまっている、つまりX > 58である割合を計算している。

## plot ECDF

Complementary Cumulative Distribution Function 


### full name, type

```{r}
p_count_full_name_type <- df_count_full_name_type |> 
  ggplot(aes(x = n, y = ecdf)) +
  geom_point() +
  scale_x_continuous(trans = "log10", breaks = 10^(0:10), name = expression(italic("x"))) +
  scale_y_continuous(trans = "log10", breaks = 10^(0:-10), name = expression(italic("Pr") ( X>= x) )) +
  labs(title = "card name + type") +
  theme_pokemon +
  theme(
    aspect.ratio = 1
  )
p_count_full_name_type
```

### full name

```{r}
p_count_full_name <-
  df_count_full_name |>
  ggplot(aes(x = n, y = ecdf)) +
  geom_point() +
  scale_x_continuous(trans = "log10", breaks = 10^(0:10), name = expression(italic("x"))) +
  scale_y_continuous(trans = "log10", breaks = 10^(0:-10), name = expression(italic("Pr") ( X>= x) )) +
  labs(title = "card name") +
  theme_pokemon +
  theme(
    aspect.ratio = 1
  )
p_count_full_name
```

### full name, remove decorated

decorated cards, like "Surfing Pikachu VMAX bababa" are omitted. Only Pikachu cards with the exact name "Pikachu" on the card are counted.

```{r}
df_count_full_name_remove_decorated |>
  ggplot(aes(x = n, y = ecdf)) +
  geom_point() +
  scale_x_continuous(trans = "log10", breaks = 10^(0:10), name = expression(italic("x"))) +
  scale_y_continuous(trans = "log10", breaks = 10^(0:-10), name = expression(italic("Pr") ( X>= x) )) +
  labs(title = "only the cards with plain pokemon name") +
  theme_pokemon +
  theme(
    aspect.ratio = 1
  )

```
・・・あんま変わらんな！！


### poke name

pokemon name. the only datatype that is comparable to the random draw null model

```{r}

df_count_pokemon |> 
 ggplot(aes(x = n, y = ecdf)) +
  geom_point() +
  scale_x_continuous(trans = "log10", breaks = 10^(0:10), name = expression(italic("x"))) +
  scale_y_continuous(trans = "log10", breaks = 10^(0:-10), name = expression(italic("Pr") ( X>= x) )) +
  labs(title = "pokemon name") +
  theme_pokemon +
  theme(
    aspect.ratio = 1
  )
```


